
---
title: "Social and configural effects on the cognitive dynamics of perspective-taking"
author: "Alexia Galati and Rick Dale"
date: "10/31/2017"
output:
  html_document: 
    keep_md: true  

  html_notebook: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())

library(pander)
library(lme4)

source('xflip.R') # this also contains some functions we'll need (print_stats)

```

## Preliminaries for Exp 1A

Here we are loading in prior data analyzed using the scripts "GDD1_batchTrajectoryAnalysis.R". 
These include data from both the "error" (control) trials and the "ambiguous" (critical) trials of the experiment. 

We trim the data by removing trials that took more than 6000 ms from the initiation of the trial to selection of an initial folder or over 1000 pixels of total distance. These values reflect long tails of the distribution, well over 3 SD of their respective distributions. We chose to omit these trials because these radical divergences from their general distribution may reflect moments when the participant is not fully engaged in the task or understanding the trial. 

We then plot a histogram for the proportion of egocentric responses across participants.

Finally, we classify participants as egocentric, other-centric, and mixed responders based on their proportion of egocentric responses on ambiguous/critical trials. For classification, we computed the proportions of egocentric and other-centric responses of each listener on critical trials, and following Duran, Dale, and Kreuz (2011), if the proportion scores exceeded .70 for one of the two perspective categories, the listener was classified as member of that category; otherwise they were classified as a mixed responder.

In Exp 1A (GDD1A), the configural cue (i.e., the orientation of the folders appearing on the tabletop) is always aligned with the perspective of the participant/listener (ego-aligned configural cue).


```{r}

load('GDD1A_churnedRawTrajectoryData.Rd')

### Data cleaning and trimming ###

# Clean by approx. 3 SDs of M for control trials
resAllError1A = resAllError[resAllError$RTDV<6000&resAllError$totalDistanceDV<1000,]
resAll1A = resAll[resAll$RTDV<6000&resAll$totalDistanceDV<1000,]

# Create an aggregate variable to see distribution of egocentrism across subjects
egoChosen = 1*(resAll1A$chosen!=resAll1A$other)
perspectiveDistribution = aggregate(egoChosen~resAll1A$fl,FUN=mean)
hist(perspectiveDistribution$egoChosen,100, main = paste("Histogram of", "proportion of egocentrism in Exp 1A"), 
    xlab='Proportion of egocentrism',ylab='Number of subjects', ylim = range(0:20))

# Construct perspective preference variables
egoSubjects = perspectiveDistribution[perspectiveDistribution$egoChosen>.7,]$resAll1A
otherSubjects = perspectiveDistribution[perspectiveDistribution$egoChosen<.3,]$resAll1A
mixedSubjects = perspectiveDistribution[(perspectiveDistribution$egoChosen>=.3 & perspectiveDistribution$egoChosen<=.7) ,]$resAll1A

length(egoSubjects)
length(otherSubjects)
length(mixedSubjects)

length(unique(resAll1A$fl))

# Label perspectivePreference levels for Ambiguous/critical trials
resAll1A$perspectivePreference = 'mixed'
resAll1A$perspectivePreference[resAll1A$fl %in% egoSubjects]='ego'
resAll1A$perspectivePreference[resAll1A$fl %in% otherSubjects]='other'

# For Control trials let's trasfer the perspectivePreference variable over 
resAllError1A$perspectivePreference = 'mixed'
resAllError1A$perspectivePreference[resAllError1A$fl %in% egoSubjects]='ego'
resAllError1A$perspectivePreference[resAllError1A$fl %in% otherSubjects]='other'

# Add label for new Experiment (Exp) variable in the dataframes for Ambiguous/critical and Control trials
resAll1A$Exp = "1A" #Experiment 1A: folder orientation is constant = always aligned with ego
resAllError1A$Exp = "1A" #Experiment 1A: folder orientation is constant = always aligned with ego

```

## Preliminaries for Exp 1B

We repeat the same process for Exp 1B.

In Exp 1B (GDD1B), the configural cue (i.e., the orientation of the folders appearing on the tabletop) is always aligned with the perspective of the task partner/speaker (other-aligned configural cue).

The two experiments are otherwise identical.

```{r}

# Same procedure for Exp 1B 
load('GDD1B_churnedRawTrajectoryData.Rd') 

### Data cleaning and trimming ###

# Clean by approx. 3 SDs of M for control trials
resAllError1B = resAllError[resAllError$RTDV<6000&resAllError$totalDistanceDV<1000,]
resAll1B = resAll[resAll$RTDV<6000&resAll$totalDistanceDV<1000,]

# Create an aggregate variable to see distribution of egocentrism across subjects
egoChosen = 1*(resAll1B$chosen!=resAll1B$other)
perspectiveDistribution = aggregate(egoChosen~resAll1B$fl,FUN=mean)

hist(perspectiveDistribution$egoChosen,100, main = paste("Histogram of", "proportion of egocentrism in Exp 1B"), 
     xlab='Proportion of egocentrism',ylab='Number of subjects', ylim = range(0:30))

# Construct perspective preference variables
egoSubjects = perspectiveDistribution[perspectiveDistribution$egoChosen>.7,]$resAll1B
otherSubjects = perspectiveDistribution[perspectiveDistribution$egoChosen<.3,]$resAll1B
mixedSubjects = perspectiveDistribution[(perspectiveDistribution$egoChosen>=.3 & perspectiveDistribution$egoChosen<=.7) ,]$resAll1B

length(egoSubjects)
length(otherSubjects)
length(mixedSubjects)

length(unique(resAll1B$fl))

# Label perspectivePreference levels for Ambiguous/critical trials
resAll1B$perspectivePreference = 'mixed'
resAll1B$perspectivePreference[resAll1B$fl %in% egoSubjects]='ego'
resAll1B$perspectivePreference[resAll1B$fl %in% otherSubjects]='other'

# For Control trials let's trasfer the perspectivePreference variable over 
resAllError1B$perspectivePreference = 'mixed'
resAllError1B$perspectivePreference[resAllError1B$fl %in% egoSubjects]='ego'
resAllError1B$perspectivePreference[resAllError1B$fl %in% otherSubjects]='other'

# Add label for new Experiment (Exp) variable in the dataframes for Ambiguous/critical and Control trials
resAll1B$Exp = "1B" #Experiment 1B: folder orientation varies = always aligned with other 
resAllError1B$Exp = "1B" #Experiment 1B: folder orientation varies = always aligned with other 

```

## Preliminaries for both experiments 

We combine the dataframes for the two experiments and recode some of the variables.

Specifically, we create an "axis" variable, for which we combine "left-right" instructions to refer to the "lateral" axis, and "front-back" instructions to refer to the "sagittal" axis. This decision is motivated in more detail in the manuscript.

We create a new variable for offset, for which we assign "90"" degree offset to speaker positions 90 and 270, as we don't expect a difference between the two. For ambiguous trials this contrasts with the "180" offset; for control trials, with "0" offset.

```{r}

# Combine the two dataframes for ambiguous/critical trials
resAll = rbind(resAll1A, resAll1B)


# Combine the two dataframes for control trials
resAllError = rbind(resAllError1A, resAllError1B)

# Let's create axis variable to compress front-back and left-right instuctions into a sagittal and lateral axis
resAll$axis = 'sagittal'
resAll$axis[resAll$instruction %in% c('righ','left')]='lateral'

# Let's compress 90 and 270 speaker positions into a 90 offset
resAll$offset = '180'
resAll$offset[resAll$ppos %in% c('init_90','init_270')]='90'

# Create egocentric choice variable
resAll$egoChoice = 1*(resAll$chosen!=resAll$other)

# Let's create the axis variable for control trials
resAllError$axis = 'sagittal'
resAllError$axis[resAllError$instruction %in% c('righ','left')]='lateral'

# Let's compress 90 and 270 speaker positions into a 90 offset
resAllError$offset = '0'
resAllError$offset[resAllError$ppos %in% c('init_90','init_270')]='90'

```

## Descriptives

Let's get some descriptives for ambiguous and control trials 

```{r}

# Ambiguous/critical trials (excluding offset, for simplicity)
pander(aggregate(egoChoice~Exp,data=resAll,FUN=mean))
pander(aggregate(RTDV~perspectivePreference+Exp+axis,data=resAll,FUN=mean))
pander(aggregate(totalDistanceDV~perspectivePreference+Exp+axis,data=resAll,FUN=mean))
pander(aggregate(xFlipDV~perspectivePreference+Exp+axis,data=resAll,FUN=mean))

# Control trials (excluding offset, for simplicity)
#pander(aggregate(err~perspectivePreference+Exp,data=resAllError,FUN=mean)) #mixed responders have high errors .25 (this is due to front-back trials, see below)
#pander(aggregate(RTDV~perspectivePreference+offset+axis+Exp,data=resAllError,FUN=mean))
#pander(aggregate(totalDistanceDV~perspectivePreference+axis+Exp,data=resAllError,FUN=mean))
#pander(aggregate(xFlipDV~perspectivePreference+axis+Exp,data=resAllError,FUN=mean))

```

## Exploratory plotting of DVs across trials to explore stabilization

We are interested in how egocentric perspective selection might differ over time in the two experiments. 
Here, we plot the proportion of egocentric choices on ambigous trials over time (i.e. across trial order).

```{r}

### Perspective choice by trial, in the two Exps ###

perspectiveByTrial1A = aggregate((1*(resAll1A$chosen!=resAll1A$other))~resAll1A$trial,FUN=mean)
perspectiveByTrial1B = aggregate(( 1*(resAll1B$chosen!=resAll1B$other))~resAll1B$trial,FUN=mean)

plot(perspectiveByTrial1A, main="Egocentric choice selection across trials in Exp 1A",
     xlab = "Trial number", ylab = "Mean proportion of egocentric choice", ylim = c(0, .90))

plot(perspectiveByTrial1B, main="Egocentric choice selection across trials in Exp 1B",
     xlab = "Trial number", ylab = "Mean proportion of egocentric choice", ylim = c(0, .90))
```

Let's look at other-centric responders only, in the two experiments. 

```{r}

perspectiveByTrial1A_Othercentric = aggregate((1*(resAll1A$chosen[resAll1A$perspectivePreference %in% c('other')]!=resAll1A$other[resAll1A$perspectivePreference %in% c('other')]))~resAll1A$trial[resAll1A$perspectivePreference %in% c('other')],FUN=mean)
perspectiveByTrial1B_Othercentric = aggregate((1*(resAll1B$chosen[resAll1B$perspectivePreference %in% c('other')]!=resAll1B$other[resAll1B$perspectivePreference %in% c('other')]))~resAll1B$trial[resAll1B$perspectivePreference %in% c('other')],FUN=mean)

plot(perspectiveByTrial1A_Othercentric, # main="Egocentric choice selection of Other-centric responders across trials in Exp 1A",
     xlab = "Trial number", ylab = "Mean proportion of egocentric choice", ylim = c(0,.90))

plot(perspectiveByTrial1B_Othercentric, #main="Egocentric choice selection of Other-centric responders across trials in Exp 1B",
     xlab = "Trial number", ylab = "Mean proportion of egocentric choice", ylim = c(0, .90))

```

## Experiment comparisons

Before getting to the LMERs, let's compare the distribution of other, ego, mixed responders across experiments.

Let's first compare the distribution of other, ego, mixed responders in Exp 1A vs. Exp 1B.

```{r}

preferenceCounts <- matrix(c(59, 18, 18, 43, 33, 17), ncol=3, byrow=TRUE)
colnames(preferenceCounts) <- c("other", "ego", "mixed")
rownames(preferenceCounts) <- c("GDD1B", "GDD1A")
preferenceCounts <- as.table(preferenceCounts)
summary(preferenceCounts)
#chisq.test(preferenceCounts)

```

Let's also compare the distributions when the classification in ego, other, mixed is based on lateral instructions only (left-right only).

This is because mixed responders (based on the previous classification, on all instruction types), made high errors on control front-back trials (esp in Exp 1B), possibly due to using a different mapping to interpret front-back. This may suggest that their responses on ambiguous trials may have not been "mixed" or random but may have simply reflected this different mapping on front-back trials. We make this point in the paper.


```{r}

preferenceCounts <- matrix(c(61, 25, 9, 29, 42, 22), ncol=3, byrow=TRUE)
colnames(preferenceCounts) <- c("other", "ego", "mixed")
rownames(preferenceCounts) <- c("GDD1B", "GDD1A")
preferenceCounts <- as.table(preferenceCounts)
summary(preferenceCounts)
#chisq.test(preferenceCounts)

```

Let's finally compare the distribution of other, ego, mixed responders in GDD1A vs. DDK study 1 
In DDK1 there were 43 other, 31 ego, and 8 mixed. 

```{r}

preferenceCounts <- matrix(c(43, 33, 17, 43, 31, 8), ncol=3, byrow=TRUE)
colnames(preferenceCounts) <- c("other", "ego", "mixed")
rownames(preferenceCounts) <- c("GDD1A", "DDKstudy1")
preferenceCounts <- as.table(preferenceCounts)
summary(preferenceCounts)
#chisq.test(preferenceCounts) 

#The two distributions don't differ significantly. Good news for replication.

```

## Linear mixed effects models for ambiguous (critical) trials

We created separate linear mixed effects models for each of the dependent variables (proprotion of egocentric choices, RT, total Distance, x-flips) on ambiguous/critical trials, combining the data from Experiment 1a and 1b. 

We start with some pre-processing of the variables, setting the reference categories where relevant, and inspecting the data structure.

```{r}

resAll = as.data.frame(as.matrix(resAll))
#Defining as factors in order to set reference categories next
resAll$Exp = as.factor(as.matrix(resAll$Exp))
resAll$offset = as.factor(as.matrix(resAll$offset))
resAll$axis = as.factor(as.matrix(resAll$axis))
resAll$perspectivePreference = as.factor(as.matrix(resAll$perspectivePreference))
resAll$trial = as.integer(as.matrix(resAll$trial))

#Setting reference categories
resAll$perspectivePreference= relevel(resAll$perspectivePreference, ref = "ego")

##Check for any "holes" in the design
#with(resAll, table(Exp, offset, axis, perspectivePreference))
#str(resAll)

#make sure DVs are of the right type
resAll$egoChoice = as.factor(as.matrix(resAll$egoChoice))
resAll$RTDV = as.numeric(as.matrix(resAll$RTDV))
resAll$totalDistanceDV = as.numeric(as.matrix(resAll$totalDistanceDV))
resAll$xFlipDV = as.integer(as.matrix(resAll$xFlipDV))

#Center time/trial
resAll$centered_trial =scale(resAll$trial)

```

## Egocentric Choice models (Ambiguous/Critical trials)

For the model with perspective choice as dependent variable (egoChoice), the fixed factors included the orientation of the folders (Exp), the speaker’s position (offset: 180° vs. 90°; within-subjects), instruction's axis (axis: on a sagittal axis: front-back vs. a lateral axis: left-right; within-subjects), and their interactions. Given that the perspective choice as a dependent variable is a binary variable, we used a logistic regression model (Jaeger, 2008).   

```{r}
ChoiceModel1 = glmer(egoChoice ~ Exp*offset*axis 
                    + (1 | fl),
                    #+ (0 + offset | fl),
                    #+ (0 + axis | fl),
                    #+ (0 + offset:axis | fl), #did not converge
                    data=resAll,
                    family = "binomial",
                    nAGQ = 1, 
                    REML = FALSE)
summary(ChoiceModel1)

#This model does not converge and is not reported in the manuscript
#ChoiceModel_time = glmer(egoChoice ~ Exp*offset*axis+ Exp*centered_trial 
                     #+ (1 | fl),
                     #+ (0 + offset | fl),
                     #+ (0 + axis | fl),
                     #+ (0 + offset:axis | fl), #did not converge
                     #data=resAll,
                     #family = "binomial",
                     # method = "Laplace",
                     #nAGQ = 1, 
                     #REML = FALSE)
#summary(ChoiceModel_time)

```

## Response Time models (Ambiguous/Critical trials)

For the models for response time (RTDV) and the other dependent measures (totalDistanceDV, xFlipDV), in addition to the fixed effects used in the egocentric choice model, we include perspective preference as a fixed effect (perspectivePreference: egocentric, other-centric, vs. mixed responder types; between-subjects), along with its interaction with the other factors.

```{r}
RTModel = lmer(log(RTDV) ~ Exp*perspectivePreference*offset*axis 
                + (1 | fl)
                + (0 + offset | fl)
                + (0 + axis | fl),
                #+ (0 + offset:axis | fl), #did not converge
                data=resAll, 
                #control = glmerControl(optimizer = 'bobyqa'),
                REML=FALSE)
print('RT:'); pander(print_stats(RTModel))

# Check residuals (before and after log transform)
# There is deviation from normality and homoscedasticity, so use log(RT)
#plot(density(resid(RTModel))) #does this look approximately normal?
#qqnorm(resid(RTModel)) #check if they fall on a straight line
#qqline(resid(RTModel)) #check departure from line

#### Investigation of stabilization across trials ####

RTModel1_time = lmer(log(RTDV) ~ Exp*perspectivePreference*offset*axis+Exp*centered_trial
                    + (1 | fl)
                    + (0 + offset | fl)
                    + (0 + axis | fl),
                    #+ (0 + centered_trial | fl),
                    #+ (0 + offset:axis | fl), 
                    data=resAll, 
                    REML=FALSE)
print('RT:'); pander(print_stats(RTModel1_time))

```

## Total Distance models (Ambiguous/Critical trials)

```{r}

DistanceModel = lmer(log(totalDistanceDV) ~ Exp*perspectivePreference*offset*axis 
                      + (1 | fl) 
                      + (0 + offset | fl)
                      + (0 + axis | fl)
                      + (0 + offset:axis | fl),
                      #+ (0 + offset:axis | fl), #did not converge
                      data=resAll, 
                      REML=FALSE) 
print('Total Distance:'); pander(print_stats(DistanceModel))

# Check residuals (before and after log transform)
# There is deviation from normality and homoscedasticity, so use log(RT)
#plot(density(resid(DistanceModel))) #does this look approximately normal?
#qqnorm(resid(DistanceModel)) #check if they fall on a straight line
#qqline(resid(DistanceModel)) #check departure from line

#### Investigation of stabilization across trials ####

DistanceModel_time = lmer(log(totalDistanceDV) ~ Exp*perspectivePreference*offset*axis+Exp*centered_trial
                    + (1 | fl)
                    + (0 + offset | fl)
                    + (0 + axis | fl),
                    #+ (0 + centered_trial | fl),
                    #+ (0 + offset:axis | fl), 
                    data=resAll, 
                    REML=FALSE)
print('Total Distance:'); pander(print_stats(DistanceModel_time))

```

## Directional Shifts models (Ambiguous/Critical trials)

```{r}
xFlipsModel = lmer(xFlipDV ~ Exp*perspectivePreference*offset*axis 
                    + (1 | fl)
                    + (0 + offset | fl)
                    + (0 + axis | fl)
                    + (0 + offset:axis | fl), 
                    data=resAll,
                    REML=FALSE)
print('Directional Shifts:'); pander(print_stats(xFlipsModel))

#### Investigation of stabilization across trials ####

xFlipsModel_time = lmer(xFlipDV ~ Exp*perspectivePreference*offset*axis+Exp*centered_trial
                          + (1 | fl)
                          + (0 + offset | fl)
                          + (0 + axis | fl),
                          #+ (0 + centered_trial | fl),
                          #+ (0 + offset:axis | fl), 
                          data=resAll, 
                          REML=FALSE)
print('Directional Shifts:'); pander(print_stats(xFlipsModel_time))

```

## Linear mixed effects models for control trials

We created separate linear mixed effects models for each of the dependent variables (proportion of errors, response time, total Distance, x-flips) on controls trials, combining the data from Experiment 1a and 1b. These models are reported in Appendix A in the manuscript.

We start with some pre-processing of the variables, setting the reference categories where relevant, and inspecting the data structure.

```{r}

resAllError = as.data.frame(as.matrix(resAllError))
#Defining as factors in order to set reference categories next
resAllError$Exp = as.factor(as.matrix(resAllError$Exp))
resAllError$offset = as.factor(as.matrix(resAllError$offset))
resAllError$axis = as.factor(as.matrix(resAllError$axis))
resAllError$perspectivePreference = as.factor(as.matrix(resAllError$perspectivePreference))

#Setting reference categories
resAllError$perspectivePreference= relevel(resAllError$perspectivePreference, ref = "ego")

##Check for any "holes" in the design
#with(resAllError, table(Exp, offset, axis, perspectivePreference))
#str(resAllError)

#make sure the DVs are of the appropriate type
resAllError$err = as.factor(as.matrix(resAllError$err))
resAllError$RTDV = as.numeric(as.matrix(resAllError$RTDV))
resAllError$totalDistanceDV = as.numeric(as.matrix(resAllError$totalDistanceDV))
resAllError$xFlipDV = as.integer(as.matrix(resAllError$xFlipDV))

```

## Proportion of errors model (Control trials)

```{r}

ErrorModel = glmer(err ~ Exp*perspectivePreference*offset*axis 
                    + (1 | fl),
                    #+ (0 + offset | fl) #does not converge
                   # + (0 + axis | fl), #does not converge
                    #+ (0 + offset:axis | fl), #does not converge
                    data=resAllError, 
                    family = "binomial",
                    REML=FALSE)
summary(ErrorModel)

```

## Response Time model (Control trials)

```{r}

RTControl = lmer(log(RTDV) ~ Exp*perspectivePreference*offset*axis 
                  + (1 | fl) 
                  + (0 + offset | fl)
                  + (0 + axis | fl)
                  + (0 + offset:axis | fl),
                  data=resAllError,  
                  REML=FALSE)
print('RT:'); pander(print_stats(RTControl))

#plot(density(resid(RTControl)))
#qqnorm(resid(RTControl))
#qqline(resid(RTControl))

```

## Total Distance models (Control trials)

```{r}

DistanceControl1 = lmer(log(totalDistanceDV) ~  Exp*perspectivePreference*offset*axis 
                        + (1 | fl) 
                        + (0 + offset | fl)
                        + (0 + axis | fl)
                        + (0 + offset:axis | fl),
                        data=resAllError,
                        REML=FALSE)
print('Total Distance:'); pander(print_stats(DistanceControl1))
summary(DistanceControl1)

#plot(density(resid(DistanceControl1)))
#qqnorm(resid(DistanceControl1))
#qqline(resid(DistanceControl1))

```

## Directional Shifts model (Control trials)

```{r}

xFlipsControl1 = lmer(xFlipDV ~ Exp*perspectivePreference*offset*axis 
                      + (1 | fl)
                      + (0 + offset | fl)
                      + (0 + axis | fl),
                      #+ (0 + offset:axis | fl), #did not converge
                      data=resAllError,
                      #family = "poisson", #models don't converge when using glmer with poisson distribution...
                      REML=FALSE)
print('Directional Shifts:'); pander(print_stats(xFlipsControl1))
summary(xFlipsControl1)

```

