
---
title: "Social and configural effects on the cognitive dynamics of perspective-taking"
authors: "Alexia Galati, Rick Dale, & Nick Duran"
date: "4/15/2017"
output:
  html_document: 
    keep_md: true  

  html_notebook: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())

library(pander)
library(lme4)
library(afex)
library(lmerTest)
library(MuMIn) 

source('xflip.R') # this also contains some functions we'll need (print_stats)

```

## Preliminaries for Exp 2

Here we are loading in prior data analyzed using the scripts "GDD2_batchTrajectoryAnalysis.R". 
These include data from both the "error" (control) trials and the "ambiguous" (critical) trials of the experiment. 

As with Exp 1A and 1B, we trim the data by removing trials that took more than 6000 ms from the initiation of the trial to selection of an initial folder or over 1000 pixels of total distance. 


In Exp 2, the configural cue (the alignment of the configuration) is manipulated within participants.


```{r}

load('GDD2_churnedRawTrajectoryData.Rd') 

### Data cleaning and trimming ###

# Clean by approx. 3 SDs of M for control trials
resAllError = resAllError[resAllError$RTDV<6000&resAllError$totalDistanceDV<1000,]
resAll = resAll[resAll$RTDV<6000&resAll$totalDistanceDV<1000,]


# Let's reverse all true/false for front/back
resAllError[resAllError$sound %in% c('fron','back'), ]$err = !resAllError[resAllError$sound %in% c('fron','back'), ]$err

```

Given the pattern of performance on sagittal (front-back) trials (with very high errors on control trials), we reverse their assignment in the data. In Exp 2, listeners appear to be interpreting front-back terms according to the opposite mapping that we had assigned. See manuscript for further explanation.


```{r}

# let's reverse front and back (since their interpretation is now the reverse of what we thought)
resAll$ego2 = 0
resAll[resAll$sound %in% c('fron','back'),]$ego2 = resAll[resAll$sound %in% c('fron','back'),]$other # new ego value with the the other one
resAll[resAll$sound %in% c('fron','back'),]$other = resAll[resAll$sound %in% c('fron','back'),]$ego
resAll[resAll$sound %in% c('fron','back'),]$ego = resAll[resAll$sound %in% c('fron','back'),]$ego2

```

As with the previous Experiments, we plot a histogram for the proportion of egocentric responses across participants. 

Again, participants are classified as egocentric, other-centric, and mixed responders based on their proportion of egocentric responses on ambiguous/critical trials, using the same criteria as Exp 1A and 1B. 


```{r}


# Create an aggregate variable to see distribution of egocentrism across subjects
#resAll$egoChosen = 1*(resAll$chosen==resAll$ego)
egoChosen = 1*(resAll$chosen==resAll$ego)
perspectiveDistribution = aggregate(egoChosen~resAll$fl,FUN=mean)
hist(perspectiveDistribution$egoChosen,100, main = paste("Histogram of", "proportion of egocentrism"), 
     xlab='Proportion of egocentrism',ylab='Number of subjects', ylim = range(0:60))


# Constructing perspective preference variables...
egoSubjects = perspectiveDistribution[perspectiveDistribution$egoChosen>.7,]$resAll
otherSubjects = perspectiveDistribution[perspectiveDistribution$egoChosen<.3,]$resAll
mixedSubjects = perspectiveDistribution[(perspectiveDistribution$egoChosen>=.3 & perspectiveDistribution$egoChosen<=.7) ,]$resAll

length(egoSubjects)
length(otherSubjects)
length(mixedSubjects)

#Label perspectivePreference levels
resAll$perspectivePreference = 'mixed'
resAll$perspectivePreference[resAll$fl %in% egoSubjects]='ego'
resAll$perspectivePreference[resAll$fl %in% otherSubjects]='other'

#for Control trials let's trasfer the perspectivePreference variable over 
resAllError$perspectivePreference = 'mixed'
resAllError$perspectivePreference[resAllError$fl %in% egoSubjects]='ego'
resAllError$perspectivePreference[resAllError$fl %in% otherSubjects]='other'

```

## Preliminaries: Variable recoding 

As with Exp 1A and 1B, we create an "axis" variable, for which we combine "left-right" instructions to refer to the "lateral" axis, and "front-back" instructions to refer to the "sagittal" axis. Again, we create an "offset" variable, for which we assign "90"" degree offset to speaker positions 90 and 270. For ambiguous trials this level contrasts with the "180" level of offset; for control trials, with the "0" offset.

Control trials also include "ego-reversed" and "partner-reversed" configurations (where the apex of the triangle is pointing toward the ego or partner, respectively); there are 2 trials for each of these types. We combine these with the "ego" and "partner" categories, respectively.

```{r}


#Let's create axis variable to compress front-back and left-right instuctions into a sagittal and lateral axis
resAll$axis = 'sagittal'
resAll$axis[resAll$sound %in% c('righ','left')]='lateral'

#Let's compress 90 and 270 speaker positions into a 90 offset
resAll$offset = '180'
resAll$offset[resAll$ppos %in% c('init_90','init_270')]='90'

#Create egocentric choice variable
resAll$egoChoice = 1*(resAll$chosen==resAll$ego)

#let's create the axis variable for control trials
resAllError$axis = 'sagittal'
resAllError$axis[resAllError$sound %in% c('righ','left')]='lateral'

#Let's compress 90 and 270 speaker positions into a 90 offset
resAllError$offset = '0'
resAllError$offset[resAllError$ppos %in% c('init_90','init_270')]='90'

#compress ego and ego-reversed configurations into a single category
#compress other and other-reversed configurations into a single category
resAllError$config[resAllError$config %in% c('ego', 'ego-reversed')]='ego'
resAllError$config[resAllError$config %in% c('partner', 'partner-reversed')]='partner'
resAllError$config = as.factor(as.matrix(resAllError$config))
#str(resAllError) #confirm 4 levels of config factor

```

## Preliminaries: Create divergence from optimal distance variable

In contrast to Exp 1A and 1B, instead of using total distance in pixels, we computed the divergence from the optimal distance to the selected folder. In the GDD2_batchTrajectoryAnalysis.R code, we computed the optimal distance to be the distance in pixels from the trial initiation to their object selection. This straight line reflects an optimal distance because it is the minimal possible distance needed to reach that point. Here, we subtract the observed distance from this optimal distance. This captured the divergence from optimal distance in pixels. 

This was done because, in the three types of configurations (ego-aligned, other-aligned, and neither-aligned), the average distance from the start point of the trial to the selected folder differed.


```{r}

resAll$divergDist = resAll$totalDistanceDV-resAll$optimalDist 
resAllError$divergDist = resAllError$totalDistanceDV-resAllError$optimalDist
#raw pixel distance from optimal

#some info about optimal distance


```

## Descriptives

Let's get some descriptives for ambiguous and control trials 

```{r}

resAll$egoChoice = as.numeric(as.matrix(resAll$egoChoice)) #if egoChoice has been converted to factor for LMERs

# Ambiguous/critical trials (excluding offset, to simplify)
pander(aggregate(egoChoice~config+axis,data=resAll,FUN=mean))
pander(aggregate(RTDV~config+perspectivePreference+axis,data=resAll,FUN=mean))
#pander(aggregate(xFlipDV~config+perspectivePreference+axis,data=resAll,FUN=mean))
#pander(aggregate(divergDist~config+perspectivePreference+axis,data=resAll,FUN=mean))

#some info about optimal distance
#summary(lm(optimalDist~egoChosen*config,data=resAll))
#pander(aggregate(optimalDist~config+egoChosen,data=resAll,FUN=mean))

#Consider the errors made by axis by responder 
#pander(aggregate(egoChoice~axis+fl+perspectivePreference,data=resAll,FUN=mean))

#This is to investigate what mixed responders are doing across trials in the two exps.
#resAllmixed = resAll[resAll$perspectivePreference == 'mixed',]
#options(max.print=999999)
#aggregate(egoChoice~trial+fl,data=resAllmixed,FUN=mean)


#resAllError$err = as.numeric(as.matrix(resAllError$err)) #if err has been converted to factor for LMERs
pander(aggregate(err~perspectivePreference+offset+axis+config,data=resAllError,FUN=mean)) 
pander(aggregate(RTDV~perspectivePreference+axis+config,data=resAllError,FUN=mean))
#pander(aggregate(xFlipDV~perspectivePreference+axis+config,data=resAllError,FUN=mean))
#pander(aggregate(divergDist~perspectivePreference+axis+config,data=resAllError,FUN=mean))

```

## Linear mixed effects models for ambiguous (critical) trials

We create separate linear mixed effects models for each of the dependent variables (proportion of egocentric choices, RT, total divergence from optimal distance, x-flips) on ambiguous/critical trials. 

We start with some pre-processing of the variables, setting the reference categories where relevant, and inspecting the data structure.

```{r}


resAll = as.data.frame(as.matrix(resAll))
#Defining as factors in order to set reference categories next
resAll$config = as.factor(as.matrix(resAll$config))
resAll$offset = as.factor(as.matrix(resAll$offset))
resAll$axis = as.factor(as.matrix(resAll$axis))
resAll$perspectivePreference = as.factor(as.matrix(resAll$perspectivePreference))
resAll$trial = as.integer(as.matrix(resAll$trial))

##Check for any "holes" in the design
with(resAll, table(config, offset, axis, perspectivePreference))
#when offset is 180, configurations cannot be ego-aligned or other-aligned, maybe remove offset as factor?
#with(resAll, table(config, axis, perspectivePreference))

#Make sure DVs are of the right type
resAll$RTDV = as.numeric(as.matrix(resAll$RTDV))
resAll$divergDist = as.numeric(as.matrix(resAll$divergDist))
resAll$xFlipDV = as.integer(as.matrix(resAll$xFlipDV))
resAll$egoChoice = as.factor(as.matrix(resAll$egoChoice))

#str(resAll)

```

## Step 1: 
Let's set up the contrast structure for each factor of interest (axis, preference, offset)

```{r}

resAll.cont <- within(resAll, {
  
  # Factor 1: instruction orientation 
  axis.latVsag <- ifelse( axis=="lateral", 1/2, 
                          ifelse( axis=="sagittal", -1/2, NA ) ) 
  
  # Factor 2: Perspective preference
  PP.othVego <- ifelse( perspectivePreference=="other", 1/2,
                        ifelse( perspectivePreference=="mixed", 0,
                                ifelse( perspectivePreference=="ego", -1/2, NA )))
  
  PP.mixVego <- ifelse( perspectivePreference=="other", 0,
                        ifelse( perspectivePreference=="mixed", 1/2,
                                ifelse( perspectivePreference=="ego", -1/2, NA )))    
  
  # Factor 3: Configuation of folders
  config.neiVego <- ifelse( config=="neither", 1/2,
                            ifelse( config=="partner", 0,
                                    ifelse( config=="ego", -1/2, NA )))
  
  config.parVego <- ifelse( config=="neither", 0,
                            ifelse( config=="partner", 1/2,
                                    ifelse( config=="ego", -1/2, NA )))
  
} )


```

##Step 2: 
Let's build the omnibus model for each DV with all effects of interest for research questions

```{r}

## Response time ##

RTModel.omni <- lmer( log(RTDV) ~ (1 | fl) + (0 + (config.neiVego + config.parVego) | fl) + (0 + axis.latVsag | fl) +
                        # perspectivePreference*config*axis
                        axis.latVsag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                      data=resAll.cont, REML=FALSE )
summary(RTModel.omni)

## Divergence from Optimal Distance ##

DistanceModel.omni <- lmer( log(divergDist) ~ (1 | fl) + (0 + (config.neiVego + config.parVego) | fl) + (0 + axis.latVsag | fl) +
                              # perspectivePreference*config*axis
                              axis.latVsag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                            data=resAll.cont, REML=FALSE )

summary(DistanceModel.omni)
#print('Distance:'); pander(print_stats(DistanceModel.omni))


## Directional shifts ##

xFlip.omni <- lmer( xFlipDV ~ axis.latVsag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego)
                    + (1 | fl),    
                    #+ (0 + (config.neiVego + config.parVego) | fl) 
                    #+ (0 + axis.latVsag | fl),
                    data=resAll.cont, 
                    #family = "poisson", 
                    REML=FALSE )
summary(xFlip.omni)
#print('Directional Shifts:'); pander(print_stats(xFlip.omni))


## Proportion of Egocentric responses ##

EgoChoiceModel.omni = glmer(egoChoice ~ (config.neiVego + config.parVego) * axis.latVsag
                     + (1 | fl)
                     #+ (0 + (config.neiVego + config.parVego) | fl) #did not converge
                     + (0 + axis.latVsag | fl), 
                     data=resAll.cont,
                     family = "binomial",
                     method = "Laplace",
                     nAGQ = 1,
                     REML = FALSE)
summary(EgoChoiceModel.omni)

```

##Step 3: 
Get the effect sizes of omnibus based on the MuMIn package - designed for mixed effects models

```{r}

##// Effect Size

r.squaredGLMM(RTModel.omni)[1]
r.squaredGLMM(RTModel.omni)[2] #we report this

r.squaredGLMM(DistanceModel.omni)[1]
r.squaredGLMM(DistanceModel.omni)[2] #we report this

r.squaredGLMM(xFlip.omni)[1]
r.squaredGLMM(xFlip.omni)[2]


r.squaredGLMM(EgoChoiceModel.omni)[1]
r.squaredGLMM(EgoChoiceModel.omni)[2]

```

##STEP 4a: 
In order to interpret the individual coefficients of the 2-way and 3-way interactions, need to first establish that the overall interactions are statistically significant
Note: Testing the 2-way interaction between perspectivePreference and config

```{r}

#This function cleans up the output of the likelihood-ratio tests
modComp = function(model1,model2,modname){
  
  llcomp = anova(model1,model2)
  df1 = llcomp$"Chi Df"[2]
  chi1 = llcomp$"Chisq"[2]
  pr1 = llcomp$"Pr(>Chisq)"[2]
  
  test1 = data.frame(c(df1,chi1,pr1))
  colnames(test1) = modname
  rownames(test1) = c("df","chi","p")
  return(test1)
}

## Note: what this does below is remove the critical 2-way interaction coefficients from the omnibus and then compares to the omni with likelihood ratio test

## what are the effects associated with the interaction between perspectivePreference and config?
# PP.othVego:config.neiVego 
# PP.othVego:config.parVego
# PP.mixVego:config.neiVego
# PP.mixVego:config.parVego

twoway.RTModel.PPvsConfig = update(RTModel.omni,.~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(RTModel.omni,twoway.RTModel.PPvsConfig,"twoway.RTModel.PPvsConfig")

twoway.DistanceModel.PPvsConfig = update(DistanceModel.omni,.~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(DistanceModel.omni,twoway.DistanceModel.PPvsConfig,"twoway.DistanceModel.PPvsConfig")

twoway.xFlipModel.PPvsConfig = update(xFlip.omni,.~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(xFlip.omni,twoway.xFlipModel.PPvsConfig,"twoway.xFlipModel.PPvsConfig")

```

##STEP 4b: 
In order to interpret the individual coefficients of the 2-way and 3-way interactions, need to first establish that the overall interactions are statisticall significant
Note: Testing the 2-way interaction between perspectivePreference and axis

```{r}

## what are the effects associated with the interaction between perspectivePreference and axis?
# axis.latVsag:PP.othVego 
# axis.latVsag:PP.mixVego

twoway.RTModel.PPvsAxis = update(RTModel.omni,.~.-axis.latVsag:PP.othVego-axis.latVsag:PP.mixVego)
modComp(RTModel.omni,twoway.RTModel.PPvsAxis,"twoway.RTModel.PPvsAxis")

twoway.DistanceModel.PPvsAxis = update(DistanceModel.omni,.~.-axis.latVsag:PP.othVego-axis.latVsag:PP.mixVego)
modComp(DistanceModel.omni,twoway.DistanceModel.PPvsAxis,"twoway.DistanceModel.PPvsAxis")

twoway.xFlip.PPvsAxis = update(xFlip.omni,.~.-axis.latVsag:PP.othVego-axis.latVsag:PP.mixVego)
modComp(xFlip.omni,twoway.xFlip.PPvsAxis,"twoway.xFlipModel.PPvsAxis")

```

##STEP 4c: 
Examining what are the effects associated with the interaction between config and axis?
axis.latVsag:config.neiVego
axis.latVsag:config.parVego 

```{r}

twoway.RTModel.ConfigvsAxis = update(RTModel.omni,.~.-axis.latVsag:config.neiVego-axis.latVsag:config.parVego)
modComp(RTModel.omni,twoway.RTModel.ConfigvsAxis,"twoway.RTModel.ConfigvsAxis")

twoway.DistanceModel.ConfigvsAxis = update(DistanceModel.omni,.~.-axis.latVsag:config.neiVego-axis.latVsag:config.parVego)
modComp(DistanceModel.omni,twoway.DistanceModel.ConfigvsAxis,"twoway.DistanceModel.ConfigvsAxis")

twoway.xFlip.ConfigvsAxis = update(xFlip.omni,.~.-axis.latVsag:config.neiVego-axis.latVsag:config.parVego)
modComp(xFlip.omni,twoway.xFlip.ConfigvsAxis,"twoway.xFlipModel.ConfigvsAxis")

```

##STEP 4d: 
In order to interpret the individual coefficients of the 2-way and 3-way interactions, need to first establish that the overall interactions are statisticall significant
Testing the 3-way interaction between perspectivePreference, config, and axis

```{r}

threeway.RTModel = update(RTModel.omni,.~.-axis.latVsag:PP.othVego:config.neiVego-axis.latVsag:PP.othVego:config.parVego-axis.latVsag:PP.mixVego:config.neiVego-axis.latVsag:PP.mixVego:config.parVego)
modComp(RTModel.omni,threeway.RTModel,"threeway.RTModel")

threeway.DistanceModel = update(DistanceModel.omni,.~.-axis.latVsag:PP.othVego:config.neiVego-axis.latVsag:PP.othVego:config.parVego-axis.latVsag:PP.mixVego:config.neiVego-axis.latVsag:PP.mixVego:config.parVego)
modComp(DistanceModel.omni,threeway.DistanceModel,"threeway.DistanceModel")

threeway.xFlipModel = update(xFlip.omni,.~.-axis.latVsag:PP.othVego:config.neiVego-axis.latVsag:PP.othVego:config.parVego-axis.latVsag:PP.mixVego:config.neiVego-axis.latVsag:PP.mixVego:config.parVego)
modComp(xFlip.omni,threeway.xFlipModel,"threeway.xFlipModel")

```

##STEP 5: 
We have an overall significant 3-way interaction, so we need to decompose holding each level of axis constant and rerun. Essentially, this is a simple effects follow-up of the 3-way

```{r}

# **Simple effects follow-up (assuming sig 3-way):** 
# Important note: Only pay attention to effects in the following models that do NOT interact with the dummy codes (axis.lat and axis.sag)...
# (i.e., ignore any fixed effect that includes "axis.lat" or "axis.sag," whether it is an interaction or main effect)"

resAll.cont <- within(resAll.cont, {
  
  axis.lat <- ifelse( axis=="lateral", 0, 
                      ifelse( axis=="sagittal", 1, NA ) ) 
  
  axis.sag <- ifelse( axis=="lateral", 1, 
                      ifelse( axis=="sagittal", 0, NA ) ) 
} )

#######
##// Axis: Lateral
#######

## RT ##

lat.omnibus.RT <- lmer( log(RTDV) ~ (1 | fl) + (0 + (config.neiVego + config.parVego) | fl) +
                          axis.lat * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                        data=resAll.cont, REML=FALSE )

##// test for two-way interaction of perspectivePreference and config in order to follow-up
lat.RT = update( lat.omnibus.RT, .~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(lat.omnibus.RT,lat.RT,"lat.omnibus.RT")


## Distance ##

lat.omnibus.Distance <- lmer( log(divergDist) ~ (1 | fl) + (0 + (config.neiVego + config.parVego) | fl) +
                                axis.lat * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                              data=resAll.cont, REML=FALSE )

##// test for two-way interaction of perspectivePreference and config in order to follow-up
lat.Distance = update( lat.omnibus.Distance, .~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(lat.omnibus.Distance,lat.Distance,"lat.omnibus.Distance")


## Directional Shifts ##

lat.omnibus.xFlip <- lmer( xFlipDV ~ (1 | fl) +
                                axis.lat * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                              data=resAll.cont, REML=FALSE )

##// test for two-way interaction of perspectivePreference and config in order to follow-up
lat.xFlip = update( lat.omnibus.xFlip, .~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(lat.omnibus.xFlip,lat.xFlip,"lat.omnibus.xFlip")



#######
##// Axis: Sagittal
#######

## RT ##

sag.omnibus.RT <- lmer( log(RTDV) ~ (1 | fl) + (0 + (config.neiVego + config.parVego) | fl) +
                          axis.sag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                        data=resAll.cont, REML=FALSE )

##// test for two-way interaction of perspectivePreference and config in order to follow-up
sag.RT = update( sag.omnibus.RT, .~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(sag.omnibus.RT,sag.RT,"sag.omnibus.RT")


## Distance ##

sag.omnibus.Distance <- lmer( log(divergDist) ~ (1 | fl) + (0 + (config.neiVego + config.parVego) | fl) +
                                axis.sag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                              data=resAll.cont, REML=FALSE )

##// test for two-way interaction of perspectivePreference and config in order to follow-up
sag.Distance = update( sag.omnibus.Distance, .~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(sag.omnibus.Distance,sag.Distance,"sag.omnibus.Distance")



summary(lat.omnibus.Distance)
#print('Lat.Distance:'); pander(print_stats(lat.omnibus.Distance))

## Directional Shifts ##

sag.omnibus.xFlip <- lmer( xFlipDV ~ (1 | fl) +
                                axis.sag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                              data=resAll.cont, REML=FALSE )

##// test for two-way interaction of perspectivePreference and config in order to follow-up
sag.xFlip = update( sag.omnibus.xFlip, .~.-PP.othVego:config.neiVego-PP.othVego:config.parVego-PP.mixVego:config.neiVego-PP.mixVego:config.parVego)
modComp(sag.omnibus.xFlip,sag.xFlip,"sag.omnibus.xFlip")


summary(lat.omnibus.xFlip)
#print('Lat.xFlip:'); pander(print_stats(lat.omnibus.xFlip))

```

## Now let's examine the effect of trial

```{r}

resAll.cont$centered_trial =scale(resAll.cont$trial)

RTModel.time <- lmer( log(RTDV) ~ (1 | fl) + 
                        (0 + (config.neiVego + config.parVego) | fl) + (0 + axis.latVsag | fl) + (0 + centered_trial | fl) +
                        centered_trial * axis.latVsag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                      data=resAll.cont, REML=FALSE )

summary(RTModel.time)
#print('RT:'); pander(print_stats(RTModel.time))

## Divergence from Optimal Distance ##

DistanceModel.time <- lmer( log(divergDist) ~ (1 | fl) +
                              (0 + (config.neiVego + config.parVego) | fl) + (0 + axis.latVsag | fl) + (0 + centered_trial | fl) +
                              centered_trial * axis.latVsag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego),        
                            data=resAll.cont, REML=FALSE )

summary(DistanceModel.time)
#print('Distance:'); pander(print_stats(DistanceModel.time))


## Directional shifts ##

xFlip.time <- lmer( xFlipDV ~ centered_trial * axis.latVsag * (PP.othVego + PP.mixVego) * (config.neiVego + config.parVego)
                    + (1 | fl)   
                    + (0 + centered_trial | fl), 
                    #+ (0 + (config.neiVego + config.parVego) | fl) 
                    #+ (0 + axis.latVsag | fl),
                    data=resAll.cont, 
                    #family = "poisson", 
                    REML=FALSE )
summary(xFlip.time)
#print('Directional Shifts:'); pander(print_stats(xFlip.time))


EgoChoiceModel.time = glmer(egoChoice ~ (config.neiVego + config.parVego) * axis.latVsag * centered_trial
                            + (1 | fl)
                            + (0 + centered_trial | fl)
                            #+ (0 + (config.neiVego + config.parVego) | fl) #did not converge
                            + (0 + axis.latVsag | fl), 
                            data=resAll.cont,
                            family = "binomial",
                            method = "Laplace",
                            nAGQ = 1,
                            REML = FALSE)
summary(EgoChoiceModel.time)


```

## Follow-up analyses excluding mixed responders (reported in Appendix B)

```{r}

resAll_NoMixed = subset(resAll, resAll$perspectivePreference %in% c('ego','other'))


resAll_NoMixed <- within(resAll_NoMixed, {
  
  # Factor 1: instruction orientation 
  axis.latVsag <- ifelse( axis=="lateral", 1/2, 
                          ifelse( axis=="sagittal", -1/2, NA ) ) 
  
  # Factor 2: Perspective preference
  PP.othVego <- ifelse( perspectivePreference=="other", 1/2,
                                ifelse( perspectivePreference=="ego", -1/2, NA ))
  
  # Factor 3: Configuation of folders
  config.neiVego <- ifelse( config=="neither", 1/2,
                            ifelse( config=="partner", 0,
                                    ifelse( config=="ego", -1/2, NA )))
  
  config.parVego <- ifelse( config=="neither", 0,
                            ifelse( config=="partner", 1/2,
                                    ifelse( config=="ego", -1/2, NA )))
  
} )


####
## STEP 2: Build the omnibus model with all effects of interest for research questions
####

## Response time ##

RTModel_NoMixed.omni <- lmer( log(RTDV) ~ (1 | fl) 
                              + (0 + (config.neiVego + config.parVego) | fl) 
                              + (0 + axis.latVsag | fl) 
                              + axis.latVsag * PP.othVego * (config.neiVego + config.parVego),        
                              data=resAll_NoMixed, REML=FALSE )
summary(RTModel_NoMixed.omni)
#print('RT:'); pander(print_stats(RTModel_NoMixed.omni))


## Divergence from Optimal Distance ##

DistanceModel_NoMixed.omni <- lmer( log(divergDist) ~ (1 | fl) 
                                    + (0 + (config.neiVego + config.parVego) | fl) 
                                    + (0 + axis.latVsag | fl) 
                                    + axis.latVsag * PP.othVego * (config.neiVego + config.parVego),        
                                    data=resAll_NoMixed, REML=FALSE )
summary(DistanceModel_NoMixed.omni)
#print('Distance:'); pander(print_stats(DistanceModel_NoMixed.omni))


## Directional shifts ##

xFlip_NoMixed.omni <- lmer( xFlipDV ~ axis.latVsag * PP.othVego * (config.neiVego + config.parVego)
                    + (1 | fl)    
                    + (0 + (config.neiVego + config.parVego) | fl) 
                    + (0 + axis.latVsag | fl),
                    data=resAll_NoMixed, 
                    #family = "poisson", 
                    REML=FALSE )
summary(xFlip_NoMixed.omni)
#print('Directional Shifts:'); pander(print_stats(xFlip_NoMixed.omni))


## Proportion of Egocentric responses ##

EgoChoiceModel_NoMixed.omni = glmer(egoChoice ~ (config.neiVego + config.parVego) * axis.latVsag
                            + (1 | fl)
                            #+ (0 + (config.neiVego + config.parVego) | fl) #did not converge
                            + (0 + axis.latVsag | fl), 
                            data=resAll_NoMixed,
                            family = "binomial",
                            method = "Laplace",
                            nAGQ = 1,
                            REML = FALSE)
summary(EgoChoiceModel_NoMixed.omni)


##// Effect Size

RTModel_NoMixed.omni <- lmer( log(RTDV) ~ (1 + (config.neiVego + config.parVego) + axis.latVsag | fl) 
                              + axis.latVsag * PP.othVego * (config.neiVego + config.parVego),        
                              data=resAll_NoMixed, REML=FALSE )
r.squaredGLMM(RTModel_NoMixed.omni)[2] #we report this


DistanceModel_NoMixed.omni <- lmer( log(divergDist) ~ (1 + (config.neiVego + config.parVego) + axis.latVsag | fl) 
                                      + axis.latVsag * PP.othVego * (config.neiVego + config.parVego),        
                                    data=resAll_NoMixed, REML=FALSE )
r.squaredGLMM(DistanceModel_NoMixed.omni)[2] #we report this


xFlip_NoMixed.omni <- lmer( xFlipDV ~ axis.latVsag * PP.othVego * (config.neiVego + config.parVego)
                            + (1 + (config.neiVego + config.parVego) + axis.latVsag | fl),
                            data=resAll_NoMixed, 
                            #family = "poisson", 
                            REML=FALSE )
r.squaredGLMM(xFlip_NoMixed.omni)[2]


EgoChoiceModel_NoMixed.omni = glmer(egoChoice ~ (config.neiVego + config.parVego) * axis.latVsag
                                    + (1 + axis.latVsag| fl), 
                                    data=resAll_NoMixed,
                                    family = "binomial",
                                    method = "Laplace",
                                    nAGQ = 1,
                                    REML = FALSE)
r.squaredGLMM(EgoChoiceModel_NoMixed.omni)[2]

```

## LMERS FOR CONTROL TRIALS (reported in Appendix A) 
Preliminaries

```{r}

resAllError = as.data.frame(as.matrix(resAllError))
#Defining as factors in order to set reference categories next
resAllError$config = as.factor(as.matrix(resAllError$config))
resAllError$offset = as.factor(as.matrix(resAllError$offset))
resAllError$axis = as.factor(as.matrix(resAllError$axis))
resAllError$perspectivePreference = as.factor(as.matrix(resAllError$perspectivePreference))


##Check for any "holes" in the design
with(resAllError, table(config, offset, axis))
#when offset is 0, configurations can only be both-aligned --> let's not include offset in the models as a factor
with(resAllError, table(config, axis, perspectivePreference))

#Make sure DVs are of the right type
resAllError$RTDV = as.numeric(as.matrix(resAllError$RTDV))
resAllError$divergDist = as.numeric(as.matrix(resAllError$divergDist))
resAllError$xFlipDV = as.integer(as.matrix(resAllError$xFlipDV))
str(resAllError)

```

##Step 1: 
Set up contrast structure

```{r}

resAllError.cont <- within(resAllError, {
  
  # Factor 1: instruction type 
  axis.latVsag <- ifelse( axis=="lateral", 1/2, 
                          ifelse( axis=="sagittal", -1/2, NA ) ) 
  
  # Factor 2: Perspective preference
  PP.othVego <- ifelse( perspectivePreference=="other", 1/2,
                        ifelse( perspectivePreference=="mixed", 0,
                                ifelse( perspectivePreference=="ego", -1/2, NA )))
  
  PP.mixVego <- ifelse( perspectivePreference=="other", 0,
                        ifelse( perspectivePreference=="mixed", 1/2,
                                ifelse( perspectivePreference=="ego", -1/2, NA )))    
  
  # Factor 3: Configuation of folders
  config.neiVego <- ifelse( config=="neither", 1/2,
                            ifelse( config=="partner", 0,
                                    ifelse( config=="both", 0,
                                      ifelse( config=="ego", -1/2, NA ))))
  
  config.parVego <- ifelse( config=="neither", 0,
                            ifelse( config=="partner", 1/2,
                                    ifelse( config=="both", 0,
                                      ifelse( config=="ego", -1/2, NA ))))
                            
  config.bothVego <- ifelse( config=="neither", 0,
                            ifelse( config=="partner", 0,
                                    ifelse( config=="both", 1/2,
                                      ifelse( config=="ego", -1/2, NA ))))
} )

```

##Step 2: 
Build the omnibus models with all effects of interest for research questions

```{r}

####Error Models#####

ErrorModel.omni = glmer(err ~ (PP.othVego+PP.mixVego) * (config.neiVego+config.parVego+config.bothVego) * axis.latVsag 
                    + (1 | fl), #does not converge!
                    #+ (0 + (config.neiVego+config.parVego+config.bothVego) | fl) #does not converge
                    #+ (0 + axis.latVsag | fl), #does not converge
                    data=resAllError.cont, 
                    family = "binomial",
                    REML=FALSE)
summary(ErrorModel.omni)


####RT Models#####

RTControl.omni = lmer(log(RTDV) ~ (PP.othVego+PP.mixVego) * (config.neiVego+config.parVego+config.bothVego) * axis.latVsag 
                  + (1 | fl) 
                  + (0 + (config.neiVego+config.parVego+config.bothVego) | fl)
                  + (0 + axis.latVsag | fl),
                  data=resAllError.cont,  
                  REML=FALSE)
#print('RT:'); pander(print_stats(RTControl.omni))
summary(RTControl.omni)

####DivergeDist models####

DistanceControl.omni = lmer(log(divergDist) ~ (PP.othVego+PP.mixVego)* (config.neiVego+config.parVego+config.bothVego) * axis.latVsag 
                      + (1 | fl) 
                      #+ (0 + (config.neiVego+config.parVego+config.bothVego) | fl) #did not converge
                      + (0 + axis.latVsag | fl),
                      data=resAllError.cont,  
                      REML=FALSE)
#print('Divergence Distance:'); pander(print_stats(DistanceControl.omni))
summary(DistanceControl.omni)

#### Directional Shifts models ####

xFlipControl.omni = lmer(xFlipDV ~ (PP.othVego+PP.mixVego) * (config.neiVego+config.parVego+config.bothVego) * axis.latVsag 
                            + (1 | fl) 
                            + (0 + (config.neiVego+config.parVego+config.bothVego) | fl) 
                            + (0 + axis.latVsag | fl),
                            data=resAllError.cont, 
                            #family = "poisson",
                            REML=FALSE)
#print('Directional Shifts:'); pander(print_stats(xFlipControl.omni))
summary(xFlipControl.omni)

```

##Step 3: 
Compute effect sizes 

```{r}

RTControl.omni = lmer(log(RTDV) ~ (PP.othVego+PP.mixVego) * (config.neiVego+config.parVego+config.bothVego) * axis.latVsag 
                      + (1 + (config.neiVego+config.parVego+config.bothVego) + axis.latVsag | fl),
                      data=resAllError.cont,  
                      REML=FALSE)

r.squaredGLMM(RTControl.omni)[2]

DistanceControl.omni = lmer(log(divergDist) ~ (PP.othVego+PP.mixVego) * (config.neiVego+config.parVego+config.bothVego) * axis.latVsag 
                            + (1 + axis.latVsag | fl),
                            data=resAllError.cont,  
                            REML=FALSE)
r.squaredGLMM(DistanceControl.omni)[2]

xFlipControl.omni = lmer(xFlipDV ~ (PP.othVego+PP.mixVego) * (config.neiVego+config.parVego+config.bothVego) * axis.latVsag 
                         + (1 + (config.neiVego+config.parVego+config.bothVego) + axis.latVsag | fl),
                         data=resAllError.cont, 
                         #family = "poisson",
                         REML=FALSE)
r.squaredGLMM(DistanceControl.omni)[2]

r.squaredGLMM(ErrorModel.omni)[2]

```

## Experiment comparisons

```{r}

###Chi-square comparison of the distribution of other, ego, mixed responders here (GDD1 -> Exp 2, in paper) vs. DDK study 1 

preferenceCounts <- matrix(c(129, 34, 21, 43, 31, 8), ncol=3, byrow=TRUE)
colnames(preferenceCounts) <- c("other", "ego", "mixed")
rownames(preferenceCounts) <- c("GDD1", "DDKstudy1")
preferenceCounts <- as.table(preferenceCounts)
summary(preferenceCounts)
chisq.test(preferenceCounts)

#GDD1 vs. GDD2a = DDK replication
#GDD2a: 43, 33, 17, other, ego, mixed
preferenceCounts <- matrix(c(129, 34, 21, 43, 33, 17), ncol=3, byrow=TRUE)
colnames(preferenceCounts) <- c("other", "ego", "mixed")
rownames(preferenceCounts) <- c("GDD1", "GDD2A")
preferenceCounts <- as.table(preferenceCounts)
summary(preferenceCounts)
chisq.test(preferenceCounts)

#GDD1 vs. GDD2b = other-aligned folders
#GDD2B: 59, 18, 18 : other, ego, mixed
preferenceCounts <- matrix(c(129, 34, 21, 59, 18, 18), ncol=3, byrow=TRUE)
colnames(preferenceCounts) <- c("other", "ego", "mixed")
rownames(preferenceCounts) <- c("GDD1", "GDD2b")
preferenceCounts <- as.table(preferenceCounts)
summary(preferenceCounts)
chisq.test(preferenceCounts)
